---
title: "Same Sex Households"
author: "Data Science"
date: "6/4/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Same Sex households in the Census

To celebrate Pride month, I was interested in looking into the Census
data for our region about the LGBTQ community.

> From
> <https://www.census.gov/content/dam/Census/library/publications/2021/acs/acsbr-005.pdf>:
> "As the largest available nationally representative household survey,
> the American Community Survey (ACS) is useful for examining the
> characteristics of households and families. Near the beginning of the
> survey questionnaire, there is a question asking how each household
> member is related to the householder. The householder is typically
> someone who owns the home or whose name is on the lease. Understanding
> each member's relationship to the householder allows us to distinguish
> various types of household composition and family structure.
>
> Definitions: Same-sex couple households: These households are
> identified using the "same-sex spouse" or "same-sex unmarried partner"
> selections to the relationship to householder question. Total coupled
> households: The total of those households in which the householder
> reports having a spouse or unmarried partner in the household.
> Includes opposite-sex and samesex married and unmarried couples.
> Unmarried partner: A person in an intimate relationship with the
> householder, such as a boyfriend or a girlfriend."

## Regional and County Totals

I wanted to look into the ACS 5-year (for good sample size) estimates
for same-sex married and unmarried couples in the region and by county.

```{r libraries, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(tidycensus)
library(writexl)
library(htmlwidgets)
source('C:/Users/SChildress/Documents/GitHub/psrc-census-tools/library/psrc_census_config.R')
source('C:/Users/SChildress/Documents/GitHub/psrc-census-tools/library/psrc_census.R')

# See: https://www.census.gov/content/dam/Census/library/publications/2021/acs/acsbr-005.pdf



Sys.getenv("CENSUS_API_KEY")

```

```{r totals}
region_same_sex_married<-psrc_acs_table("B09019_011", "region", 2019,'acs5')

region_same_sex_cohabit<-psrc_acs_table("B09019_013", "region", 2019,'acs5')

# Regional Households

region_hhs<- psrc_acs_table("B08201_001", "region", 2019,'acs5')

# I should actually retrieve these values from the table above (lazy)

same_sex_share <-(12259+7311)/1603060

same_sex_share

county_same_sex_married<-psrc_acs_table("B09019_011", "county", 2019,'acs5')
county_same_sex_cohabit<-psrc_acs_table("B09019_013", "county", 2019,'acs5')
```

The data says that there are 12,300 same-sex married couples and 7,300
same-sex unmarried couples. This just seems way too low at around 1.2%
of households. There are obviously many same-sex people who are living
without partners, but regardless, it makes me question the data.

I retrieved the same sex households by Census Tract to see what the
geographic results look like. Many tracts are within the MOE.

## Same Sex Couples by Census Tract

```{r maps}
tract_same_sex_married<-psrc_acs_table("B09019_011", "tract", 2019,'acs5')
tract_same_sex_cohabit<-psrc_acs_table("B09019_013", "tract", 2019,'acs5')


ss_married_map<-create_tract_map(tract_same_sex_married)
ss_cohabit_map<-create_tract_map(tract_same_sex_cohabit)

ss_married_map
ss_cohabit_map
```

I guess the maps look reasonable?
